### 2021-11-06 publication/experiments-detailed.tex
12-pages

Table~\ref{tab:results} summarises the results for each experiment. We also apply the 
dimensionality reduction method 
of Section~\ref{sec:dim-reduction}
to the two higher dimensional datasets (the school and FB egonet). 
For this we use equation~(\ref{eqn:c-star}) with $k=1$,
in order to reduce the dimension from 
$D$ to a desired $D'$. 
We then retrain the feature-block predictor using only the retained 
feature set $\Dcal'$, and report the log-loss over the training and 
test sets for the reduced classifier -- 
denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. 
These values are also given in Table~\ref{tab:results}.

Based on these results, some remarks are in order.
Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher 
than the training loss $\bar{\Lcal}_0$. This is expected,
as the test set is smaller than the training set and hence 
more susceptible to variability in its construction. 
Secondly, it can be seen that the dimensionality reduction procedure 
brings the training and test losses closer together. This implies that 
the features we keep are indeed correlated with the underlying graphical 
partition and that the approach generalises correctly.

The average description length per entity,
$\bar{S}_e$, of the graph, 
has very small variance, suggesting that
the detected communities can be found reliably (to within an arbitrary 
relabelling of blocks). For reference, we plot an inferred partition for each 
of the graphs in Figure~\ref{fig:graphs-all}. The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at dividing the graph into densely connected clusters.

\subsubsection{Political books}

We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component.  
Furthermore, in Table~\ref{tab:results} we see that the training and test losses 
are very similar and both are low in magnitude. This is strong evidence 
that political affiliation is a very appropriate explanatory 
variable for the overall network structure.

However, from Figure~\ref{fig:polbooks-accuracy} we see that block 1 has low accuracy. 
This suggests that detected block 1 is not solely composed of ``neutral" books but also 
contains some ``liberal" and ``conservative" authors. Examining 
Figure~\ref{fig:polbooks-graph}, we see the majority of paths between blocks 2 and 3 go through block 1.
Block 1 is in effect a bridge between the ``conservative'' and ``liberal'' blocks so it is unsurprising that some books from either side leak into block 1.

\subsubsection{Primary school dynamic contacts}

We choose the number of communities $B=10$, in line with the total number of 
school classes. From Figure~\ref{fig:school-null}, we see only the pupils' class memberships (1A-5B) have survived
the dimensionality-reduction process for $D'=10$;
gender and teacher/student status have been discarded,
meaning these are poor predictors of overall macro-structure.

The vast majority of blocks are composed of a single class. 
However, some blocks have two comparably strong classes as their predictors. 
For example, blocks 2 and 5 both contain classes 3A and 3B as their 2 best predictors. 
This suggests that the social divide between classes is less pronounced 
for pupils in year 3. Conversely, some classes are found to extend over two 
detected blocks (class 2B spans blocks 8 and 9) but we do 
not have a feature which explains the division.

Figure~\ref{fig:school-accuracy} shows excellent accuracy for the majority of blocks. In fact the only blocks with low accuracy are those that have a school-class span two blocks such that we cannot reliably distinguish between the two. This is more pronounced when we apply hard classification rather than the soft cross-entropy loss. Indeed block 1 has low accuracy because class 1A has a higher weight component for block 3 than it does for block 1. As we use hard classification to compute accuracy, vertices belonging to class 1A are predicted to belong to block 3. A similar effect is seen in block 5 which often loses out to block 2. The same can be said of blocks 8 and 9. However, in this case the weights for class 2B are very similar -- thus explaining the roughly evenly split accuracy for blocks 8 and 9.

\subsubsection{Facebook egonet}

We choose $B=10$ and $D'=10$ for this experiment. The selected features 
(Figure~\ref{fig:fb-null}) are those that best explain the high-level 
community structure. The majority of them are education related. 
Nevertheless, for $D'=10$ we only have good explanations for some of the detected blocks; several blocks in 
Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$. This is further emphasised by the disparate accuracies in Figure~\ref{fig:fb-accuracy}. Nevertheless, observe that the accuracy is only high for blocks that contain high magnitude weights. The only exception to this rule is block 9 which nonetheless may have high magnitude weights just below the cut-off $c^*$.

When the feature dimension is very large, it becomes increasingly likely that 
a particular feature may uniquely identify a small set of nodes. If these nodes 
are all part of the same community, then the classifier may overfit for that 
particular parameter. The regularisation term imposed by the prior goes some 
way towards alleviating this problem. Nevertheless, we see in 
Figure~\ref{fig:fb-null} that the feature \verb*|birthday-5| has a very 
high weight as it relates to block 1 -- but it is highly unlikely
that birthdays determine graphical structure.

>>>>

For the higher-dimensional datasets, we also apply the 
dimensionality reduction method 
of Section~\ref{sec:dim-reduction}.  
We then retrain the feature-block predictor using only the retained 
feature set $\Dcal'$, and report the log-loss over the training and 
test sets for the reduced classifier -- 
denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively.

Table~\ref{tab:results} summarises the results for each experiment.
We see that the dimensionality reduction procedure 
brings the training and test losses closer together. This implies that 
the features we keep are indeed correlated with the underlying graphical 
partition and that the approach generalises correctly.
%
The average description length per entity,
$\bar{S}_e$, of the graph, 
has very small variance, suggesting that
the detected communities can be found reliably (to within an arbitrary 
relabelling of blocks).

\paragraph{\textbf{Political books.}}

We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component.  
Furthermore, the training and test losses from Table~\ref{tab:results}  
are very similar and both are low in magnitude. This is strong evidence 
that political affiliation is a very appropriate explanatory 
variable for the overall network structure.
%
However, from Figure~\ref{fig:polbooks-accuracy} we see that block 1 has low accuracy. 
This suggests that detected block 1 is not solely composed of ``neutral" books but also 
contains some ``liberal" and ``conservative" authors. Examining 
Figure~\ref{fig:polbooks-graph}, we see the majority of paths between blocks 2 and 3 go through block 1.
Block 1 is in effect a bridge between the ``conservative'' and ``liberal'' blocks so it is unsurprising that some books from either side leak into block 1.

\paragraph{\textbf{Primary school.}}

We choose the number of communities $B=10$, in line with the total number of 
school classes. Only the pupils' class memberships (1A-5B) survive
the dimensionality-reduction process (Figur~\ref{fig:school-null}e);
gender and teacher/student status have been discarded,
meaning these are poor predictors of overall macro-structure.
%
The vast majority of blocks are composed of a single class. 
However, some blocks have two comparably strong classes as their predictors (e.g. blocks 2 and 5). 
Conversely, some classes are found to extend over two 
detected blocks (class 2B spans blocks 8 and 9) but we do 
not have a feature which explains the division.
%
Figure~\ref{fig:school-accuracy} shows excellent accuracy for the majority of blocks. In fact the only blocks with low accuracy are those that have a school-class span two blocks such that we cannot reliably distinguish between the two. This is more pronounced when we apply hard classification rather than the soft cross-entropy loss. Perhapsthere are unobserved features which explain this divide.

\paragraph{\textbf{Facebook egonet.}}

We choose $B=10$ and $D'=10$ for this experiment. The selected features 
(Figure~\ref{fig:fb-null}) are those that best explain the high-level 
community structure. The majority of them are education related. 
Nevertheless, for $D'=10$ we only have good explanations for some of the detected blocks; several blocks in 
Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$. This is further emphasised by the disparate accuracies in Figure~\ref{fig:fb-accuracy}.

### 2021-11-06 publication/experiments.tex
12-pages

We apply the outlined methods to a variety of datasets:

@@

Next, to assess the performance of the feature-to-block predictor, 
we randomly partition the vertex set $[N]$ into training and test sets: $\Gcal_0$ and $\Gcal_1$.
The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain. As $|\Gcal_0| \neq |\Gcal_1|$ in general, we use the average cross-entropy loss 
over each set to gauge fit,
%
\begin{equation}
	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star^{(t)},
	\quad \textrm{where} \quad
	\Lcal_\star^{(t)} \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
	\label{eqn:cross-entropy-loss}
\end{equation}
%
where $\star \in \{0, 1\}$ toggles between the training and test sets.
%
Nevertheless, the cross-entropy loss is a coarse measure of fit. We wish to define a new measure of fit specific to each detected block. Let us define
$
	\Bcal_\star(j) \coloneqq \{i \in \Gcal_\star : \hat{b}_i = j\}
$
where
$ 
	\hat{b}_i \coloneqq \argmax_j \hat{y}_{ij}.
$
Now $\Bcal_\star(j)$ is the set of vertices that have maximum a posteriori probability of belonging to block $j$. We now define the accuracy for block $j$ as,
%
\begin{equation}
	\eta_\star(j) \coloneqq \frac{1}{|\Bcal_\star (j)| \cdot 
	|\Tcal_\theta| } 
	\sum_{i \in \Bcal_\star (j)}  \sum_{t \in \Tcal_\theta}
	\one \left\{\hat{b}_i = \argmax_j \phi_j \left( x_i; \theta^{(t)} \right) \right\}.
	\label{eqn:accuracy}
\end{equation}

This effectively tests whether the feature-to-block and the graph-to-block predictions agree in their largest component. We call this metric, $\eta_\star(j)$, the {\em block-accuracy} for block $j$. It is clearly bounded $0 \leq \eta_\star(j) \leq 1$, with an accuracy of 1 meaning perfect agreement for the vertices in detected block $j$.

\input{experiments-detailed}
\input{experiments-figures}

>>>>

We apply the outlined methods to a variety of labelled networks. For reference, the inferred partitions for all of these are potted on Figure~\ref{fig:graphs-all}.

@@

Next, to assess the performance of the feature-to-block predictor, 
we randomly partition the vertex set $[N]$ into training and test sets: $\Gcal_0$ and $\Gcal_1$.
\footnote{We choose $f=0.7$ as the fraction of vertices that go to form our test set. Please refer to \url{https://github.com/LozzaTray/Jormungandr-code} for full source code.}
The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain. As $|\Gcal_0| \neq |\Gcal_1|$ in general, we use the average cross-entropy loss 
over each set to gauge fit,
%
\begin{equation}
	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star^{(t)},
	\quad \textrm{where} \quad
	\Lcal_\star^{(t)} \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
	\label{eqn:cross-entropy-loss}
\end{equation}
%
where $\star \in \{0, 1\}$ toggles between the training and test sets.
%
Nevertheless, the cross-entropy loss is a coarse measure of fit. We wish to define a new measure of fit specific to each detected block. Let us define
$
	\Bcal_\star(j) \coloneqq \{i \in \Gcal_\star : \hat{b}_i = j\}
$
where
$ 
	\hat{b}_i \coloneqq \argmax_j \hat{y}_{ij}.
$
Now $\Bcal_\star(j)$ is the set of vertices that have maximum a posteriori probability of belonging to block $j$. We now define the {\em block-accuracy} for block $j$ as in \ref{eqn:accuracy}. This effectively tests whether the feature-to-block and the graph-to-block predictions agree in their largest component.
%
\begin{equation}
	\eta_\star(j) \coloneqq \frac{1}{|\Bcal_\star (j)| \cdot 
	|\Tcal_\theta| } 
	\sum_{i \in \Bcal_\star (j)}  \sum_{t \in \Tcal_\theta}
	\one \left\{\hat{b}_i = \argmax_j \phi_j \left( x_i; \theta^{(t)} \right) \right\}.
	\label{eqn:accuracy}
\end{equation}

\input{experiments-figures}
\input{experiments-detailed}
